### 2018-05-26 abstract.tex
Fix spelling mistakes using spelling checker

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  are utilized to effectively exploit the capability of on-chip caches.
  These techniques are sufficient for lastly-recent-used (LRU) caches,
  which are often used in high-performance
  processors deployed in high-end servers and supercomputers.
  But in recent years the market has become increasingly more diverse
  in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to GEMM development in a multi-threaded context.
  As several threads try to load data into the shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in GEMM routines,
  by partitioning the shared cache into physically disjoint
  sets and assign different sets to different threads.
  We implement SCP in the OpenBLAS library and
  evaluate the performance on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared (by every 4 cores) psudo-random L2 caches.
  The results show that conflict misses on both L1 and L2
  caches are effectively reduced, resulting in an improvement
  of the GEMM performance by $2.75\%$--$6.91\%$.
\end{abstract}

>>>>

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  are utilized to effectively exploit the capability of on-chip caches.
  These techniques are sufficient for lastly-recent-used (LRU) caches,
  which are often used in high-performance
  processors deployed in high-end servers and supercomputers.
  But in recent years the market has become increasingly more diverse
  in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to GEMM development in a multi-threaded context.
  As several threads try to load data into the shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in GEMM routines,
  by partitioning the shared cache into physically disjoint
  sets and assign different sets to different threads.
  We implement SCP in the OpenBLAS library and
  evaluate the performance on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared (by every 4 cores) pseudo-random L2 caches.
  The results show that conflict misses on both L1 and L2
  caches are effectively reduced, resulting in an improvement
  of the GEMM performance by $2.75\%$--$6.91\%$.
\end{abstract}

### 2018-05-26 background.tex
Fix spelling mistakes using spelling checker

\subsection{The Memory Hierachy Review}\label{subsec:hierachy}

Modern architectures take advantage of fast on-chip caches
to overcome the ``memory wall''.
As a compromise between direct mapped cache and full-associative cache,
set-associative cache has been the dominant choice for real-world architectures.
To understand how cache sharing can cause inter-thread cache conflicts
and hurts GEMM performance, we review some details
about the working mechanisim of set-associative caches and
the memory hierachy of modern processors.

@@

It is easy to prove that any continous memory block no larger than $c/n$
cannot conflict with itself.
We define a quantity, way-capacity, denoted as $wc=c/n$,
to represent this size bound for conflict-free memory blocks.

The main memory, $NL$ levels of caches, and $NT$ processor cores
form a memory hierachy generally organized into
a $NL+2$ level regular tree structure.
Processor cores at level $0$ are leaves,
caches at levels $1$ -- $NL$ are intermedia nodes
and the main memory at level $NL+1$ is the tree root.
Any node in the memory hierachy can be identified by
its layer and its unique index on the layer $(layer, index)$.
The number of nodes on layer $l$ is $NT / nt_l$.

Throughout this paper we will use a 4-core processor with
$NL=2$ levels of caches as the platform for SCP demonstration.
Figure.~\ref{fig:hierachy} shows its memory hierachy and
detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
Each core has 32 128-bit (16B) vector registers,
each capable of storing 2 double precision floating-point numbers.
For convenience, a processor core on level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with programmable replacement policy.

\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{figures/cluster-new}
  \caption{Memory hierachy of the 4-core processor
    (nodes labeled with $(layer,index)$)}
  \label{fig:hierachy}
\end{figure}

\begin{table}
  \centering
  \caption{Architectural parameters of the 4-core processor}
  %% (see Section~\ref{subsec:hierachy} for meanings of symbols in the header row)
  \label{tab:cluster}
  \begin{tabular}{lccccccl}
    \toprule
    type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & policy \\
    \midrule
    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & programmable \\
    L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & psudo-random \\
    \bottomrule
  \end{tabular}
\end{table}

@@

In this blocked algorithm, the loops over the $N$, $K$ and $M$
matrix dimensions are tiled by sizes $N_c$, $K_c$ and $M_c$ at 
layers 1, 2 and 3, respectively.
At layers 4 and 5, the $N$ and $M$ dimensions are further tiled by sizes
$N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ submatrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 does not have to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices on each level fit into a certain level in the memory hierachy.
Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
the constraints on these tiling factors.
$es$ denotes the size of matrix element, e.g. 8B for double precision floating-point numbers,
and $c_l$ denotes the capacity of registers or caches on level $l$.
First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
(the psudo level-0 cache).
Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
at layer 5 fit into the L1 cache.
Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
at layer 4 fit into the L2 cache.
At last, by (\ref{eq:constraints.l3}), $N_c$ is constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
Larger matrices on higher layers 1 -- 3 always reside in main memory.

@@

\begin{comment}
Table~\ref{tab:factors} shows the procedure choosing various tiling factors.
In each row, one or more tiling factors (column 1) is determined
by a size constraint (column 3) related to a certain level
in the memory hierachy (column 2).
\begin{table}
  \centering
  \caption{Procedure of choosing tiling factors}
  \label{tab:factors}
  \begin{tabular}{cccl}
    \toprule
    factor & memory & constraint \\
    \midrule
    $M_r$, $N_r$ & registers & $M_r + N_r + M_r N_r \le Regs$ \\
    $K_c$        & L1 cache  & $N_r K_c + 2 M_r K_c \le L1$ \\
    $M_c$        & L2 cache  & $M_c K_c + 2 N_r K_c \le L2$ \\
    $N_c$        & L3 cache  & $N_c K_c + M_c K_c \le L3$ \\
    \bottomrule
  \end{tabular}
\end{table}
\end{comment}

Goto~\cite{gotogemm} factors out the innermost three loops at layers 4 -- 6 for
computing $C_2\ = C_2 + \alpha A_2 B_2$ as an architecture-dependent kernel,
known as  GEBP (GEneral multiply of a Block of $A$ and a Panel of $B$).
It is worthy noting that before GEBP is called (at layer 3),
$A_1[ii:ii+M_c-1][:]$ and $B_1$ are packed into $A_2$ and $B_2$
in a special continous layout,
so consecutive memory access is ensured within the GEBP kernel.
GotoBLAS~\cite{gotoblas}, and its successor, OpenBLAS~\cite{openblas},
implement GEMM programs based on this factorization,
with GEBP highly optimized for the target processor (often coded in assembly).

As stated in Section~\ref{sec:intro}, a GEMM implementation
is made up by a kernel routine and an overall strategy.
In Figure.~\ref{fig:gemm}, GEBP surves as the kernel routine,
and the overall strategy is defined by the three outter loops at layers 1 -- 3.
Developers can control how to partition the workload by choosing
tile factors $M_c$, $N_c$ and $K_c$,
and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
Either the N loop at layer 1 or M loop at layer 3, or both of them,
can be parallelized in a multi-threaded context.
Specifically, the implementation in Figure.~\ref{fig:gemm}
chooses a NKM loop order and parallelizes the M loop at layer 3.
Different threads work on different parts of $A_1$
and share the same $B_1$.
As a result, there are multiple $A_2$ instances, one per thread,
and a single $B_2$ instances, shared by all threads.
To understand how the loop on layer 3 is parallelized on the 4-core processor, 
Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
While all color shaded submatrices are accessed by $T_1$,
only z-curve masked submatrices (a subset of color shaded ones)
are packed by the thread.
That is, each thread packs its own $A_2$ instance and 
$B_2$ is packed by all thread collabratively.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

@@

In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar in that it is expected to
reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
In general, even with carefully selected tile factors,
matrix data can still be evicted out for a few reasons.
For instance, data structures other than the packed matrices,
or program code (in case of a unified cache) may be fetched into a
cache set the matrix data lives in.
We call this kind of cache conflicts intra-thread cache conflicts
because it occurs within the execution of a single thread.
To deal with intra-thread conflicts,
the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to L1 cache
in every iteration at layer 6 before it actually reads them in case of
the data being evicted out accidentally.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefething strategy
can reduce the penalty caused by intra-thread conflicts.

In contrary to intra-thread cache conflicts,
there exists another kind of cache conflicts,
inter-thread cache conflicts, on architectures with shared caches.
Suppose two working threads, $T_0$ and $T_1$, are running independently
on the 4-core processor in Figure.~\ref{fig:hierachy}
to accomplish a GEMM operation.
At time $t_s$, $T_0$ prefetches some data at address $addr_0$
to cache set $\varphi_2(addr_0)$ of the shared L2 cache.
The data will be actually read at time $t_e$.
Then at some time $t_m$ ($t_s < t_m < t_e$),
$T_1$ prefetches data at addr $addr_1$ to L2 cache.
If $\varphi_2(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,
data at $addr_1$ and $addr_0$ will be fetched to the same cache set.
As the shared L2 cache has a non-LRU policy,
data at $addr_0$ may suffer an eviction before it is actually read,
i.e. a conflict miss occurs.

>>>>

\subsection{The Memory Hierarchy Review}\label{subsec:hierachy}

Modern architectures take advantage of fast on-chip caches
to overcome the ``memory wall''.
As a compromise between direct mapped cache and full-associative cache,
set-associative cache has been the dominant choice for real-world architectures.
To understand how cache sharing can cause inter-thread cache conflicts
and hurts GEMM performance, we review some details
about the working mechanism of set-associative caches and
the memory hierarchy of modern processors.

@@

It is easy to prove that any continuous memory block no larger than $c/n$
cannot conflict with itself.
We define a quantity, way-capacity, denoted as $wc=c/n$,
to represent this size bound for conflict-free memory blocks.

The main memory, $NL$ levels of caches, and $NT$ processor cores
form a memory hierarchy generally organized into
a $NL+2$ level regular tree structure.
Processor cores at level $0$ are leaves,
caches at levels $1$ -- $NL$ are intermediate nodes
and the main memory at level $NL+1$ is the tree root.
Any node in the memory hierarchy can be identified by
its layer and its unique index on the layer $(layer, index)$.
The number of nodes on layer $l$ is $NT / nt_l$.

Throughout this paper we will use a 4-core processor with
$NL=2$ levels of caches as the platform for SCP demonstration.
Figure.~\ref{fig:hierachy} shows its memory hierarchy and
detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
Each core has 32 128-bit (16B) vector registers,
each capable of storing 2 double precision floating-point numbers.
For convenience, a processor core on level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with programmable replacement policy.

\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{figures/cluster-new}
  \caption{Memory hierarchy of the 4-core processor
    (nodes labeled with $(layer,index)$)}
  \label{fig:hierachy}
\end{figure}

\begin{table}
  \centering
  \caption{Architectural parameters of the 4-core processor}
  %% (see Section~\ref{subsec:hierachy} for meanings of symbols in the header row)
  \label{tab:cluster}
  \begin{tabular}{lccccccl}
    \toprule
    type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & policy \\
    \midrule
    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & programmable \\
    L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & pseudo-random \\
    \bottomrule
  \end{tabular}
\end{table}

@@

In this blocked algorithm, the loops over the $N$, $K$ and $M$
matrix dimensions are tiled by sizes $N_c$, $K_c$ and $M_c$ at 
layers 1, 2 and 3, respectively.
At layers 4 and 5, the $N$ and $M$ dimensions are further tiled by sizes
$N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ sub-matrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 does not have to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices on each level fit into a certain level in the memory hierarchy.
Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
the constraints on these tiling factors.
$es$ denotes the size of matrix element, e.g. 8B for double precision floating-point numbers,
and $c_l$ denotes the capacity of registers or caches on level $l$.
First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
(the pseudo level-0 cache).
Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
at layer 5 fit into the L1 cache.
Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
at layer 4 fit into the L2 cache.
At last, by (\ref{eq:constraints.l3}), $N_c$ is constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
Larger matrices on higher layers 1 -- 3 always reside in main memory.

@@

\begin{comment}
Table~\ref{tab:factors} shows the procedure choosing various tiling factors.
In each row, one or more tiling factors (column 1) is determined
by a size constraint (column 3) related to a certain level
in the memory hierarchy (column 2).
\begin{table}
  \centering
  \caption{Procedure of choosing tiling factors}
  \label{tab:factors}
  \begin{tabular}{cccl}
    \toprule
    factor & memory & constraint \\
    \midrule
    $M_r$, $N_r$ & registers & $M_r + N_r + M_r N_r \le Regs$ \\
    $K_c$        & L1 cache  & $N_r K_c + 2 M_r K_c \le L1$ \\
    $M_c$        & L2 cache  & $M_c K_c + 2 N_r K_c \le L2$ \\
    $N_c$        & L3 cache  & $N_c K_c + M_c K_c \le L3$ \\
    \bottomrule
  \end{tabular}
\end{table}
\end{comment}

Goto~\cite{gotogemm} factors out the innermost three loops at layers 4 -- 6 for
computing $C_2\ = C_2 + \alpha A_2 B_2$ as an architecture-dependent kernel,
known as  GEBP (GEneral multiply of a Block of $A$ and a Panel of $B$).
It is worthy noting that before GEBP is called (at layer 3),
$A_1[ii:ii+M_c-1][:]$ and $B_1$ are packed into $A_2$ and $B_2$
in a special continuous layout,
so consecutive memory access is ensured within the GEBP kernel.
GotoBLAS~\cite{gotoblas}, and its successor, OpenBLAS~\cite{openblas},
implement GEMM programs based on this factorization,
with GEBP highly optimized for the target processor (often coded in assembly).

As stated in Section~\ref{sec:intro}, a GEMM implementation
is made up by a kernel routine and an overall strategy.
In Figure.~\ref{fig:gemm}, GEBP serves as the kernel routine,
and the overall strategy is defined by the three outer loops at layers 1 -- 3.
Developers can control how to partition the workload by choosing
tile factors $M_c$, $N_c$ and $K_c$,
and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
Either the N loop at layer 1 or M loop at layer 3, or both of them,
can be parallelized in a multi-threaded context.
Specifically, the implementation in Figure.~\ref{fig:gemm}
chooses a NKM loop order and parallelizes the M loop at layer 3.
Different threads work on different parts of $A_1$
and share the same $B_1$.
As a result, there are multiple $A_2$ instances, one per thread,
and a single $B_2$ instances, shared by all threads.
To understand how the loop on layer 3 is parallelized on the 4-core processor, 
Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
While all color shaded sub-matrices are accessed by $T_1$,
only z-curve masked sub-matrices (a subset of color shaded ones)
are packed by the thread.
That is, each thread packs its own $A_2$ instance and 
$B_2$ is packed by all thread collaboratively.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

@@

In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar in that it is expected to
reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
In general, even with carefully selected tile factors,
matrix data can still be evicted out for a few reasons.
For instance, data structures other than the packed matrices,
or program code (in case of a unified cache) may be fetched into a
cache set the matrix data lives in.
We call this kind of cache conflicts intra-thread cache conflicts
because it occurs within the execution of a single thread.
To deal with intra-thread conflicts,
the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to L1 cache
in every iteration at layer 6 before it actually reads them in case of
the data being evicted out accidentally.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefetching strategy
can reduce the penalty caused by intra-thread conflicts.

In contrary to intra-thread cache conflicts,
there exists another kind of cache conflicts,
inter-thread cache conflicts, on architectures with shared caches.
Suppose two working threads, $T_0$ and $T_1$, are running independently
on the 4-core processor in Figure.~\ref{fig:hierachy}
to accomplish a GEMM operation.
At time $t_s$, $T_0$ prefetches some data at address $addr_0$
to cache set $\varphi_2(addr_0)$ of the shared L2 cache.
The data will be actually read at time $t_e$.
Then at some time $t_m$ ($t_s < t_m < t_e$),
$T_1$ prefetches data at address $addr_1$ to L2 cache.
If $\varphi_2(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,
data at $addr_1$ and $addr_0$ will be fetched to the same cache set.
As the shared L2 cache has a non-LRU policy,
data at $addr_0$ may suffer an eviction before it is actually read,
i.e. a conflict miss occurs.

### 2018-05-26 evaluation.tex
Fix spelling mistakes using spelling checker

As the L2 cache is shared by cores inside the cluster,
the impact of cache sharing can be demonstrated
by comparing DGEMM performance with the same $NT$
but different $NT_C$s.
Figure.~\ref{fig:drawback} shows the evaluation results.
Each curve represents a parallelism configuration.
Different subfigures are configured with different $NT$s,
and different curves in the same subfigure are
configured with different $NT_C$s.
As there is only one configuration for $NT=64$ ($NC=16$, $NT_C=4$)
and no comparation is available,
results of $NT=64$ is combined with $NT=32$ in Figure.~\ref{fig:drawback-32}.

@@

We evaluate SCP using the same method as in Section~\ref{subsec:drawback}.
SCP is not evaluated with $NT_C=1$ because there is no need
to partition the cache if $NT_C=1$.
Figure.~\ref{fig:benefit} shows the evaluation results.
For the purpose of comparation, results without SCP
(denoted as Base in the figure) are also presented,

@@

To understand why SCP outperforms Base,
and whether SCP actually eliminate inter-thread conflicts,
we use the PAPI~\cite{papi} profiling tool to
analyze the cache behavior of the GEMM program.
Four cache related hardware events are counted,
which are listed and briefly described in Table.~\ref{tab:events}.
With these hardware events,
miss rates of L1 and L2 caches can be calculated with
$\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
The hardware events are mesured in a per-thread manner
and all presented results are the average over threads.

@@

Privatization inccurs more packing overhead because
redundant packing of $B_2$ is enforced.
Figure.~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelism configuration $NT_C=4$ and $NC=4$.
The whole packing workload of $B_2$ is divided into 16 tasks
and the packing workload of thread $T_1$ is masked by z-curves. 
%% and the task assignment is annoted in Figure.~\ref{fig:packb}.
In conventional packing, each thread takes a single task.
In full privatization, all tasks are replicated to all threads.
As a result, the packing overhead grows in proportion to $NT$
in full privatization strategy, which is unacceptable
if GEMM is highly parallelized, e.g. $NT=64$.
Partial privatization in Figure.~\ref{fig:packb.partial}
offers an alternative to full privatization.
In partial privatization, privatization only occurs inside
a cluster so that tasks belong to one cluster need not to be
replicated to other clusters.
Consequently, extra packing overhead of the partial privatization
strategy is bounded by $NT_C$.
For instance, on Phytium 2000+, overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in conventional packing.

Figure.~\ref{fig:privb} shows the evaluation results of the partial
privatization strategy with $NT_C=4$ and $NC=4$.
Results under other parallelism configurations are similar.
For the purpose of comparation,
SCP with and without privatization are both presented
(denoted as SCP-P and SCP in Figure.~\ref{fig:privb}, respectively).
Figure.~\ref{fig:privb.papi} shows the cache miss rates.
Privatization achieves slightly lower miss rates on both L1 and L2 caches.
Figure.~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower cache miss rates,
privatization suffers a performance lose compared to SCP.
Why performance drops while cache miss rates get reduced?
The reason lies in the extra packing overhead introduced by privatization.
Figure.~\ref{fig:privb.breakdown} shows the occupancy ratio
of overall executing time of various overheads.
Overheads in GEMM includes
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While SCP-P involves roughly the same overheads in
packing $A_2$ and synchronization as SCP,
it spends much more time in packing $B_2$.
Table.~\ref{tab:breakdown} lists the breakdown of
executing time. We can see that overhead of packing $B_2$
is multiplied by a factor of $NT_C=4$,
and the total overhead increases from $2.35\%$ to $4.72\%$.

@@

Results in this section demonstrats that
privatization of the shared matrix $B_2$ can further
eliminate inter-thread cache conflicts.
But the improvement in cache performance is very small because
$B_2$ only occupies a small portion of the shared L2 cache,
and this marginal improvement is not sufficient to
balance the extra packing overhead introduced
by privatization.

>>>>

As the L2 cache is shared by cores inside the cluster,
the impact of cache sharing can be demonstrated
by comparing DGEMM performance with the same $NT$
but different $NT_C$s.
Figure.~\ref{fig:drawback} shows the evaluation results.
Each curve represents a parallelism configuration.
Different sub-figures are configured with different $NT$s,
and different curves in the same sub-figure are
configured with different $NT_C$s.
As there is only one configuration for $NT=64$ ($NC=16$, $NT_C=4$)
and no comparison is available,
results of $NT=64$ is combined with $NT=32$ in Figure.~\ref{fig:drawback-32}.

@@

We evaluate SCP using the same method as in Section~\ref{subsec:drawback}.
SCP is not evaluated with $NT_C=1$ because there is no need
to partition the cache if $NT_C=1$.
Figure.~\ref{fig:benefit} shows the evaluation results.
For the purpose of comparison, results without SCP
(denoted as Base in the figure) are also presented,

@@

To understand why SCP outperforms Base,
and whether SCP actually eliminate inter-thread conflicts,
we use the PAPI~\cite{papi} profiling tool to
analyze the cache behavior of the GEMM program.
Four cache related hardware events are counted,
which are listed and briefly described in Table.~\ref{tab:events}.
With these hardware events,
miss rates of L1 and L2 caches can be calculated with
$\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
The hardware events are measured in a per-thread manner
and all presented results are the average over threads.

@@

Privatization introduces more packing overhead because
redundant packing of $B_2$ is enforced.
Figure.~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelism configuration $NT_C=4$ and $NC=4$.
The whole packing workload of $B_2$ is divided into 16 tasks
and the packing workload of thread $T_1$ is masked by z-curves. 
%% and the task assignment is annoted in Figure.~\ref{fig:packb}.
In conventional packing, each thread takes a single task.
In full privatization, all tasks are replicated to all threads.
As a result, the packing overhead grows in proportion to $NT$
in full privatization strategy, which is unacceptable
if GEMM is highly parallelized, e.g. $NT=64$.
Partial privatization in Figure.~\ref{fig:packb.partial}
offers an alternative to full privatization.
In partial privatization, privatization only occurs inside
a cluster so that tasks belong to one cluster need not to be
replicated to other clusters.
Consequently, extra packing overhead of the partial privatization
strategy is bounded by $NT_C$.
For instance, on Phytium 2000+, overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in conventional packing.

Figure.~\ref{fig:privb} shows the evaluation results of the partial
privatization strategy with $NT_C=4$ and $NC=4$.
Results under other parallelism configurations are similar.
For the purpose of comparison,
SCP with and without privatization are both presented
(denoted as SCP-P and SCP in Figure.~\ref{fig:privb}, respectively).
Figure.~\ref{fig:privb.papi} shows the cache miss rates.
Privatization achieves slightly lower miss rates on both L1 and L2 caches.
Figure.~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower cache miss rates,
privatization suffers a performance lose compared to SCP.
Why performance drops while cache miss rates get reduced?
The reason lies in the extra packing overhead introduced by privatization.
Figure.~\ref{fig:privb.breakdown} shows the occupancy ratio
of overall executing time of various overheads.
Overheads in GEMM includes
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While SCP-P involves roughly the same overheads in
packing $A_2$ and synchronization as SCP,
it spends much more time in packing $B_2$.
Table.~\ref{tab:breakdown} lists the breakdown of
executing time. We can see that overhead of packing $B_2$
is multiplied by a factor of $NT_C=4$,
and the total overhead increases from $2.35\%$ to $4.72\%$.

@@

Results in this section demonstrates that
privatization of the shared matrix $B_2$ can further
eliminate inter-thread cache conflicts.
But the improvement in cache performance is very small because
$B_2$ only occupies a small portion of the shared L2 cache,
and this marginal improvement is not sufficient to
balance the extra packing overhead introduced
by privatization.
